Arterial pulsations are not powerful enough to drive perivascular drainage of solutes from the brain

### 2017-03-13 paper.tex
experimental results added

@@

\title{Arterial pulsations are not powerful enough to drive perivascular drainage of solutes from the brain}
\author[1,2]{Alexandra K. Diem}
\author[2]{Neil W. Bressloff}
\author[3]{Roxana O. Carare}
\author[4]{Giles Richardson}
\affil[1]{Institute for Complex Systems Simulation, University of Southampton, Southampton SO17 1BJ, UK}
\affil[2]{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO17 1BJ, UK}
\affil[3]{Faculty of Medicine, University of Southampton, Southampton, UK}
\affil[4]{Faculty of Mathematical Sciences, University of Southampton, Southampton, UK}

@@

A widely accepted hypothesis for the driving mechanism of this reverse perivascular drainage of solutes from the brain is that it is driven by arterial pulsations \cite{Weller2009,Carare2008,Hawkes2011,Morris2014,Schley2006,Attems2011,Wang2011,Iliff2012,Asgari2015,Sharp2015,Weller2015a}. In this paper we tested this hypothesis by using a multi-scale modelling approach combining analytical and numerical methods. The BM is modelled as a porous medium using Darcy's law and a lubrication approximation. Flow inside the BM is driven by both the ISF pressure gradient and the deformation of the arterial wall, which are obtained from a numerical model of the middle cerebral artery (MCA). To guarantee reverse drainage (\ie in the opposite direction of the blood flow) a valve mechanism is implemented as a pressure gradient dependent permeablility. It is shown that this valve mechanism is neccesary to achieve net reverse drainage. The results show that this valve mechanism successfully ensures net reverse drainage inside the BM and is neccesary. However, due to the long wavelength of arterial pulsations the pressure gradients inside the BM were found to be too small to drive fluid flow at the expected velocities. The resulting velocity of the model was approximately two orders of magnitude smaller than values obtained experimentally \cite{Carare2008}. Therefore arterial pulsations are concluded to be insufficient to drive perivascular drainage along the BM of cerebral arteries.

\section{Porous medium model of lymphatic drainage in the BM}
\label{sec:diodemodel}

@@

\section{Numerical Model of a Straight Arterial Section}

@@

\section{Perivascular Drainage through the MCA}

@@

@@

The results show that the valve mechanism as implemented in our analytical model of the BM guarantees net reverse flow inside the BM under physiological blood flow conditions. The valve mechanism is required to be strong, \ie the ratio of $K_0/K_1$ needs to be small. Additionally, the model shows that a valve mechanism is neccessary for net reverse flow to occur within the BM, confirming results from previous studies that did not include a valve mechanism and concluded the existence of some attachment mechanism \cite{Schley2006,Wang2011}. Net drainage additionally depends on the distance of the BM from the artery lumen. The closer the BM is placed to the lumen the faster net reverse drainage is, but variations remain within the same order of magnitude. This is due to the relatively large thickness of cerebral artery walls and their elastic properties. The hypothesis that the BM utilises a unidirectional valve mechanism is plausible as the general lymphatic system of the body also has one and its existence has previously been speculated \cite{Weller2010,Schley2006,Heppell2013}.

>>>>

\DeclareSIUnit\Molar{\textsc{M}}

@@

\title{Arterial pulsations are not powerful enough to drive perivascular drainage of solutes from the brain}
\author[1,2]{Alexandra K. Diem}
\author[2]{Neil W. Bressloff}
\author[3]{Roxana O. Carare}
\author[3]{Matthew MacGregor Sharp}
\author[4]{Giles Richardson}
\affil[1]{Institute for Complex Systems Simulation, University of Southampton, Southampton SO17 1BJ, UK}
\affil[2]{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO17 1BJ, UK}
\affil[3]{Faculty of Medicine, University of Southampton, Southampton, UK}
\affil[4]{Faculty of Mathematical Sciences, University of Southampton, Southampton, UK}

@@

A widely accepted hypothesis for the driving mechanism of this reverse perivascular drainage of solutes from the brain is that it is driven by arterial pulsations \cite{Weller2009,Carare2008,Hawkes2011,Morris2014,Schley2006,Attems2011,Wang2011,Iliff2012,Asgari2015,Sharp2015,Weller2015a}. In this paper we tested this hypothesis by using a multi-scale modelling approach combining analytical and numerical methods. The BM is modelled as a porous medium using Darcy's law and a lubrication approximation. Flow inside the BM is driven by both the ISF pressure gradient and the deformation of the arterial wall, which are obtained from a numerical model of the middle cerebral artery (MCA). To guarantee reverse drainage (\ie in the opposite direction of the blood flow) a valve mechanism is implemented as a pressure gradient dependent permeablility. It is shown that this valve mechanism is neccesary to achieve net reverse drainage. The results show that this valve mechanism successfully ensures net reverse drainage inside the BM and is neccesary. However, due to the long wavelength of arterial pulsations the pressure gradients inside the BM were found to be too small to drive fluid flow at the expected velocities. The resulting velocity of the model was approximately two orders of magnitude smaller than values obtained experimentally \cite{Carare2008}. Therefore arterial pulsations are concluded to be insufficient to drive perivascular drainage along the BM of cerebral arteries. The simulation results are confirmed by an experimental assessment of the effect of pulse modulation in a mouse model, which showed no significant difference in perivascular drainage compare to control mice.

\section{Numerical modelling of perivascular drainage of \Ab}

First, the velocity of perivascular drainage in a pressure-pulse driven scenario is evaluated using computational modelling. An analytical model of the BM was developed, which captures the preference of flow in the reverse direction to the blood flow. Pressure inside the BM is evaluated using a numerical model of blood pressure inside the artery and a stress analysis in the artery wall. These were then used to evaulate the velocity of perivascular drainage through the MCA.

\subsection{Porous medium model of lymphatic drainage through the BM}
\label{sec:diodemodel}

@@

\subsection{Numerical Model of a Straight Arterial Section}

@@

\subsection{Perivascular Drainage through the MCA}

@@

\section{Experimental assessment of intramural perivascular drainage of \Ab}

In addition to performing simulations to assess the feasibility of blood-pressure driven arterial pulsations providing the main driving mechanism for perivascular drainage through the BM of the cerebral vasculature, mouse model experiments were performed to confirm the results \textit{in vivo}.

\subsection{Intracerebral injections}

All mice were kept on a standard 12-hr light/dark cycle and allowed food and water \textit{ad libitum}. All experiments were carried out in accordance with animal care guidelines stipulated by the Animal Care and Use Committee at the University of Southampton and the Home Office (PPL 20/2095).

To minimise the effects of anaesthesia on mouse cardiovascular function, isoflurane was chosen as the anaesthetizing agent. To reduce arterial pulsations, the beta-blocker Atenolol (Ate, \SI{10}{\milli\gram\per\kilo\gram}) was administered intraperitoneally (IP). Mouse vital parameters were monitored throughout the procedure using a MouseOX Plus and an infrared mouse thigh sensor (STARR Life Sciences, Holliston, MA, USA). Thermal regulation was maintained using a heat pad and rectal probe (BASi, West Lafayette, IN, USA). All intracerebral injections were performed using a \SIrange{1}{5}{\micro\litre} Hirschmann microcapillary pipette (Sigma UK) with a tip adjusted to a diameter of $< \SI{50}{\micro\metre}$ using a Sutter P97 Flaming Brown Pipette puller.

Ten week old wild-type C57B16 mice ($n = 4$) were anaesthetized with isoflurane, placed prone on a stereotaxic instrument and secured with head adaptors. Baseline measurements of oxygen saturation, heart rate, breath rate and pulse distention were taken before administering Ate. A volume of \SI{0.5}{\micro\litre} of \Abs HiLyte Fluor 555 (Cambridge Bioscience, UK) was stereotaxically injected into the left hippocampus at a rate of \SI{0.2}{\micro\litre\per\minute} (coordinates from Bregma: $\textrm{AP} = \SI{-1.9}{\milli\metre}$, $\textrm{ML} = \SI{1.5}{\milli\metre}$ and $\textrm{DV} = \SI{-1.7}{\milli\metre}$) \SI{10}{\minute} post Ate injection. Injection pipettes were left \textit{in situ} for \SI{2}{\minute} to prevent reflux. Mice were euthanized \SI{5}{\minute} after withdrawal of the injection pipette through overdose with pento-barbital (\SI{200}{\milli\gram\per\kilo\gram}). Mice were then intracardially perfused with \SI{0.01}{\Molar} phosphate buffered saline (PBS), pH~7.2 followed by \SI{4}{\percent} paraformaldehyde in \SI{0.01}{\Molar} PBS, pH~7.2. Brains were removed, post fixed overnight in fresh fixative and then processed for immunohistochemistry. For controls, ten week old wild-type C57B16 ($n = 3$) were processed following the same method but without administration of Ate.

\subsection{Immunohistochemistry}

Brains were cut into \SI{20}{\micro\metre} coronal sections using a freezing microtome (Leica) and the injection site located by examining on a Zeiss Axioskop 2 fitted with a rhodamine filter. Sections of \SI{400}{\micro\metre} and \SI{800}{\micro\metre} thickness both posterior and anterior to the injection site were selected for immunohistochemistry. Sections were blocked in \SI{15}{\percent} goat serum (Sigma 9023) for \SI{1}{\hour} at room temperature (RT). They were then incubated in primary antibodies diluted in \SI{0.01}{\Molar} phosphate buffered saline \SI{0.1}{\percent} triton x100 (PBSt). Primary antibodies comprised rabbit anti-collagen IV (AbCam, ab6586) 1/400 and mouse anti-smooth muscle actin (SMA) FITC conjugated (Sigma F3777) 1/200. Sections were incubated overnight at \SI{4}{\celsius} in a moist chamber. Following washing 3x\SI{10}{\minute} \SI{0.01}{\Molar} PBS sections were incubated in goat anti-rabbit Alexa Fluor$\textregistered$ 633 (ThermoFisher Scientific A-21070) 1/200 PBSt.

\subsection{Image analysis}

Images of the whole hippocampus were captured on a Leica SP8 confocal microscope fitted with LASX software in the tile scanning mode with sequential imaging. Once captured, images were exported as TIFF files and uploaded into Adobe Photoshop CS6. An area of hippocampus was drawn extending transversely from CA3 pyramidal cells to the apex of the blades of the granule cell layer and from the ventricle to \textit{stratum oriens} in the superior/inferior axis. \Ab containing vessels were defined according to size and presence of SMA: capillaries had a diameter of $\leq \SI{10}{\micro\metre}$, veins had no SMA and a diameter of $> \SI{10}{\micro\metre}$. The number of vessels was expressed per \SI{}{\square\milli\metre} of hippocampus.

Statistical analysis was performed by an independent t-test, with \SI{95}{\percent} set as confidence interval.

\subsection{Results}

\begin{figure}
  \centerline{\includegraphics{figures/confocal.jpg}}
  \caption{Caption\label{fig:confocal}}
\end{figure}

\begin{table}
  \center
  \caption{Caption\label{tab:caption}}
  \begin{tabular}{lcccccc}
    \toprule
     & \multicolumn{3}{c}{Atenolol} & \multicolumn{3}{c}{Saline}\\
    \midrule
    arteries    & 6.19 & 0.66 & 4 & 7.48  & 2.4  & 3\\
    veins       & 0.79 & 0.3  & 4 & 0.3   & 0.15 & 3\\
    capillaries & 0.52 & 0.14 & 4 & 0.6   & 0.3  & 3\\
    \midrule
    arteries    & 6.79 & 1.4  & 4 & 15.03 & 5.26 & 3\\
    veins       & 0.62 & 0.05 & 4 & 1.8   & 1.08 & 3\\
    capillaries & 0.17 & 0.1  & 4 & 2.34  & 1.55 & 3\\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}
  \centerline{\includegraphics{figures/graph.png}}
  \caption{Caption\label{fig:graph}}
\end{figure}

Within \SI{5}{\minute} of injections into the hippocampus, fluorescent \Ab was observed diffusely in the hippocampus and co-localizing with collagen IV in the walls of capillaries and arteries, with very few veins involved in both atenolol-treated and control mice (Figure~\ref{fig:caption} and Table~\ref{tab:confocal}). Application of an independent t-test to the data revealed no significant differences between the saline and atenolol treated mice either anterior or posterior to the injection site (Figure~\ref{fig:graph}).

@@

The results show that the valve mechanism as implemented in our analytical model of the BM guarantees net reverse flow inside the BM under physiological blood flow conditions. The valve mechanism is required to be strong, \ie the ratio of $K_0/K_1$ needs to be small. Additionally,the model shows that a valve mechanism is neccessary for net reverse flow to occur within the BM, confirming results from previous studies that did not include a valve mechanism and concluded the existence of some attachment mechanism \cite{Schley2006,Wang2011}. Net drainage additionally depends on the distance of the BM from the artery lumen. The closer the BM is placed to the lumen the faster net reverse drainage is, but variations remain within the same order of magnitude. This is due to the relatively large thickness of cerebral artery walls and their elastic properties. The hypothesis that the BM utilises a unidirectional valve mechanism is plausible as the general lymphatic system of the body also has one and its existence has previously been speculated \cite{Weller2010,Schley2006,Heppell2013}.
